---
title: "1. Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cleaningkit)
```

## Overview

The `cleaningkit` package is built to automate the process of survey data cleaning and validation. The workflow of this package is divided into two primary phases:

1. **The Validation and Cleaning Process**
2. **Applying the Cleaning Logs to Produce the Final Output**

### 1. The Validation and Cleaning Process

This initial phase involves running a suite of validation functions (all starting with `validate_`). Each of these functions performs a specific check on your dataset and generates a corresponding log of flagged issues. 

After running the individual checks, you will combine all the generated logs into a single, comprehensive cleaning log that also includes a dataset to facilitate the review process. Concurrently, you will extract, prepare, and create a specialized cleaning log for handling "other" text responses.

For detailed instructions and examples on running these functions, please see the [Validation and Cleaning Process](2-validation-and-cleaning.html) article.

### 2. Applying the Cleaning Logs

Once the cleaning logs (including the "other" text responses log) have been manually reviewed and filled out, you proceed to the application phase. 

This phase involves:

* Reading the reviewed cleaning logs back into R.
* Evaluating the logs to ensure they match the raw data correctly.
* Applying the specified cleaning actions to the dataset.
* Reviewing the cleaned data to verify that all the required cleaning steps have been successfully implemented.
* Creating a final consolidated output workbook. This final output contains the README, the raw data, the newly cleaned data, and the applied cleaning logs all in one place.

For more details on this step, please see the [Applying Cleaning Logs](3-applying-cleaning-logs.html) article.
