---
title: "2. Data Cleaning and Validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2. Data Cleaning and Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Initialize the package and all required packages

```{r}
rm(list = ls())
library(cleaningkit)

cleaningkit::load_packages()
```

## Read tool data

```{r}
tool_survey <- cleaningkit::read_tool_survey("./resources/tool.xlsx")
tool_choices <- cleaningkit::read_tool_choices("./resources/tool.xlsx")
```

## Create other response db

```{r}
other_labels <- cleaningkit::get_other_labels(
  tool_survey = tool_survey,
  other_text_types = c("Q31_2")
)

other_db <- cleaningkit::get_other_db(
  other_labels = other_labels,
  tool_choices = tool_choices,
  tool_survey = tool_survey
)
```

## Read raw data

```{r}
raw_data <- cleaningkit::read_raw_data(
  filename = "./data/data.xlsx",
  tool_survey = tool_survey
)
```

## Filter read raw data

Perform a filter on the raw dataset for multiple cleaning logs outputs.

## Prepare and save other responses

Add the other responses questions you want to be included in the output in the question argument.

```{r}
df <- cleaningkit::prepare_other_responses(
  uuid_column = "_uuid",
  raw_data = raw_data,
  other_db = other_db,
  tool_choices = tool_choices,
  extra_columns = c("username"),
  questions = c(
    "Q34_1",
    "Q35_1",
    "Q37_1",
    "Q38_1",
    "Q39_1",
    "Q41_1",
    "Q33_1"
  )
)

cleaningkit::save_other_responses(
  df = df,
  other_db = other_db,
  save_location = "./output/other",
  enumerator_id = "username"
)
```

## Validate duration

Flags anything below 15 mins and above 60 mins.

```{r}
duration_log <- cleaningkit::validate_duration(
  dataset = raw_data,
  column_to_check = "_duration",
  uuid_column = "_uuid",
  log_name = "duration_log",
  lower_bound = 15,
  upper_bound = 60,
  skip_label_row = TRUE
)
```

## Validate completeness

`metadata_cols = c("start","end","today","deviceid","username","phonenumber","_uuid")` has a list of all metadata columns that will be ingored during this check.

```{r}
completeness_log <- cleaningkit::validate_completeness(
  dataset = raw_data,
  uuid_column = "_uuid",
  log_name = "completeness_log",
  min_content_cells = 100,
  skip_label_row = TRUE
)
```

## Validate refused

Flags any surveys that have more than 6 "Refused" by default.

```{r}
refused_log <- cleaningkit::validate_refused(
  dataset = raw_data,
  uuid_column = "_uuid",
  log_name = "refused_log",
  max_refused = 6,
  refused_value = "Refused",
  skip_label_row = TRUE
)
```

## Validate back to back interviews

Flags any interviews from the same enumerator if they are 10 mins apart by default.

```{r}
back_to_back_log <- cleaningkit::validate_back_to_back(
  dataset = raw_data,
  uuid_column = "_uuid",
  enumerator_column = "username",
  start_column = "start",
  end_column = "end",
  log_name = "back_to_back_log",
  threshold_hours = 0,
  threshold_mins = 10,
  gap_from = c("end", "start"),
  skip_label_row = TRUE
)
```

## Validate country of interview

Flags if the respondent is being interviewed in their country of nationality.

```{r}
country_of_interview_log <- cleaningkit::validate_country_of_interview(
  dataset = raw_data,
  uuid_column = "_uuid",
  country_interview_col = "Q13",
  nationality_col = "Q31",
  journey_start_col = "Q41",
  log_name = "country_of_interview_log",
  skip_label_row = TRUE
)
```

## Validate time of interview

Flags the earliest and latest times of interview. By default earliest hour is 5am and latest hour is 10pm.

```{r}
interview_time_log <- cleaningkit::validate_interview_time(
  dataset = raw_data,
  uuid_column = "_uuid",
  time_column = "start",
  log_name = "interview_time_log",
  earliest_hour = 5,
  latest_hour = 22,
  flag_missing = FALSE,
  skip_label_row = TRUE
)
```

## Validate duplicate surveys

Groups data by enumerator and checks surveys which are similar.

```{r}
duplicate_log <- raw_data %>%
  cleaningkit::validate_duplicates(
    tool_survey = tool_survey,
    idnk_value = "Don't know",
    threshold = 30
  )
```

## Validate duplicate questions

Groups data by enumerator and checks questions passed for similarity.

```{r}
duplicate_questions_log <- raw_data %>%
  cleaningkit::validate_duplicate_questions(
    questions_to_check = c("Q161_1", "Q162_1", "Q152_1", "P2P18_1")
  )
```

## Validate logical

Reads the logical excel sheets and uses that for validating the survey.

```{r}
logical_list <- openxlsx::read.xlsx("./resources/logical_checks_mmc.xlsx", sheet = 1)

logical_check_log <- raw_data %>%
  cleaningkit::validate_logical_with_list(
    list_of_check = logical_list,
    check_id_column = "check_id",
    check_to_perform_column = "check_to_perform",
    columns_to_clean_column = "columns_to_clean",
    description_column = "description"
  )
```

## Combine logs

```{r}
list_of_log_all <- c(
  duration_log,
  completeness_log,
  refused_log,
  back_to_back_log,
  country_of_interview_log,
  interview_time_log,
  logical_check_log,
  duplicate_log,
  duplicate_questions_log
)

combined_log <- cleaningkit::create_combined_log(
  list_of_log = list_of_log_all,
  dataset_name = "checked_dataset"
)
```

## Create cleaning log

```{r}
cleaningkit::create_cleaning_log(
  write_list = combined_log,
  output_path = paste0(
    "output/",
    Sys.Date(),
    "_follow-ups.xlsx"
  )
)
```
